Introduction of using AWS instance to run experiments

### 2022-03-30 docs/run_experiments.Rmd
A few updates related to do prediction

According to some experiments, the accuracy of cropland is relatively low. One reason is that the farming condition over Tanzania is complex and thus hard to model. Another reason is the validation of this class has relatively low quality as well. There are limited things we could do for these reasons. But we can let the model to consider more about cropland when training. The strategy can be increasing the loss weight of cropland class. However, the effect is not conclusive.

#### Ensemble training

When training a full model using the whole dataset, the main problem is cropland. One reason is that
cropland highly relies on local climate and cultivating conditions and Tanzania is a large country.

- Train multiple models with subset of labels.

  - Zone 1, 2, 3, 4, 5 (composite learning rate)
  - Zone 1, 2, 3, 4
  - Zone 1, 2, 3, 5
  - Zone 1, 2, 4, 5
  - Zone 1, 3, 4, 5
  - Zone 2, 3, 4, 5

- Ensemble these models by voting, possibly at segment level.

Records:

Zone 1, 2, 3, 4, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_full_12bs_4augs_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 151
```
Zone 1, 2, 3, 4
```
python hrlcm/fit_compose_lr.py --exp_name unet_full_12bs_4augs_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 151
```
Zone 1, 2, 3, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_full_12bs_4augs_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 151
```
Zone 1, 2, 4, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_full_12bs_4augs_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 151
```
Zone 1, 3, 4, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_full_12bs_4augs_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 151
```
Zone 2, 3, 4, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_full_12bs_4augs_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 151
```

@@

@@

For example:
```
python hrlcm/predict.py --args_path 'results/dl/unet_zone1_12bs_4augs_80epochs/checkpoints/args.pkl' --checkpoint_file 'results/dl/unet_zone1_12bs_4augs_80epochs/checkpoints/final.pth' --fname_predict 'dl_catalog_test.csv' --data_dir 'results/tanzania' --out_dir 'results/prediction' --stats_dir 'results/tanzania/norm_stats' --num_workers 48 --batch_size 32 --gpu_devices '0,1,2,3'

scp -r ubuntu@ec2-54-163-146-125.compute-1.amazonaws.com:~/hrlcm/results/prediction /Users/pinot/Dropbox/research/hrlcm/results/tanzania/prediction/unet_zone1_12bs_4augs_80epochs
```

@@

>>>>

According to some previous experiments, the accuracy of cropland is relatively low. One reason is that the farming condition over Tanzania is complex and thus hard to model. Another reason is the validation of this class has relatively low quality as well. There are limited things we could do for these reasons. But we can let the model to consider more about cropland when training. The strategy can be increasing the loss weight of cropland class. However, the effect is not conclusive.

Hum... Using a bigger weight for cropland in loss calculation increased the overall accuracy a bit. However, the cost is the overestimation of cropland over sparse vegetation area.

#### Freeze layers

Sometimes, people freeze some layers, only update weights of last few layers. The parameters for U-Net used in this project are:

1 * 8 (initial) + 4 * 8 (down) + 4 * 8 (up) + 2 = 74 layers

- 40: freeze all down layers
- 56: freeze all down and the first two up blocks
- 64: freeze all down and the first three up blocks
- 72: freeze all down and up blocks, just update the last two layers

#### U-Net structure

Using `ConvTranspose2d` instead of `Upsample` could reduce the overestimation issue of built-up areas. Because upsampling just uses bilinear sampling method, which could smooth the whole image. However, built-up objects are usually tiny. Thus, they are affected by bilinear upsampling a lot.

However, there are also cons of `ConvTranspose2d`. Using `ConvTranspose2d` loses the advantage of bilinear smoothing. Thus, at the boundary of an object, it might be some salt noises. If post-segmentation pruning is applied, this is not a big issue.

Another issue is that because using `ConvTranspose2d` means add another learnable parameter layer, it slightly increases the need of computational resources. It really depends if this is an issue.

Thus, whether using `ConvTranspose2d` or not depends what you want.

#### "Hierachical" training

When training a full model using the whole dataset, the main problem is cropland. One reason is that
cropland highly relies on local climate and cultivating conditions and Tanzania is a large country. To reduce this impact, we could shrink the study area gradually into each agro-ecological zone.

- Modeling structure:

```
1, 2, 4, 5 -> 1, 2, 5 -> 1, 2 -> 1
                      -> 2, 5 -> 2
                              -> 5
2, 3, 4, 5 -> 2, 3, 4 -> 2, 3 -> 3
                      -> 2, 4 -> 4
```

Zone 1, 2, 3, 4, 5

Comment out line 211 and add `learning_rates = [args.max_lr] * args.epochs` to use a fixed learning rate 0.001 to train.

```
python hrlcm/fit_compose_lr.py --exp_name unet_full_startup_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --max_lr 0.001 --save_freq 20 --epochs 81
```

Then, discard the changes above.

Zone 1, 2, 4, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_z1245_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 121 --resume 'results/dl/unet_full_startup_80epochs/checkpoints/final.pth'
```

Comment out line 249 in `fit_compose_lr.py`, and modify `lr_scheduler.py` correspondingly. The learning rate scheduler is saved as a pickle as well.

Zone 1, 2, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_z125_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.0001 --save_freq 10 --epochs 81 --resume 'results/dl/unet_z1245_120epochs/checkpoints/final.pth'

# Fine tune using a smaller learning rate
python hrlcm/fit_compose_lr.py --exp_name unet_z125_ft_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.000001 --max_lr 0.00001 --save_freq 10 --epochs 81 --resume 'results/dl/unet_z125_80epochs/checkpoints/final.pth'
```

Discard the changes above.

Zone 2, 3, 4, 5
```
python hrlcm/fit_compose_lr.py --exp_name unet_z2345_120epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.001 --save_freq 20 --epochs 121 --resume 'results/dl/unet_full_startup_80epochs/checkpoints/final.pth'
```

Do the same as zone 1, 2, 5.

Zone 2, 3, 4
```
python hrlcm/fit_compose_lr.py --exp_name unet_z234_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.0001 --save_freq 20 --epochs 81 --resume 'results/dl/unet_z2345_120epochs/checkpoints/final.pth'

# Fine tune with a smaller learning rate
python hrlcm/fit_compose_lr.py --exp_name unet_z234_ft_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.000001 --max_lr 0.00001 --save_freq 10 --epochs 81 --resume 'results/dl/unet_z234_80epochs/checkpoints/final.pth'
```

**[NOTE:]** _I tried multiple ways, which includes adding more training labels, using different types of loss functions (particularly ones to deal with class imbalance), and applying different model structure. No luck to get a reasonable balanced results if shrink the training area further into single zones. So I decided to stop here to use three zones as a group for the final result._

_But if you like, or I get the right method to do further, go further to finish the following steps._

_Zone 1_
```
# "Shrink" the model to zone1
python hrlcm/fit_compose_lr.py --exp_name unet_z125_ew_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.0001 --save_freq 10 --epochs 81 --resume 'results/dl/unet_z125_80epochs/checkpoints/final.pth'

# "Fine-tune" the zone1 model
python hrlcm/fit_compose_lr.py --exp_name unet_z1_ft_72fz_cl_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --max_lr 0.00000001 --save_freq 10 --epochs 81 --resume 'results/dl/unet_z1_ew80epochs/checkpoints/final.pth'
```

_Zone 2, 5 (For zone 2 & 5)_
```
python hrlcm/fit_compose_lr.py --exp_name unet_z25_ew_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.00001 --max_lr 0.0001 --save_freq 20 --epochs 81 --resume 'results/dl/unet_z125_80epochs/checkpoints/final.pth'
```

_Zone 2, 3 (For zone 3)_
```
python hrlcm/fit_compose_lr.py --exp_name unet_z23_12bs_4augs_ft_fz56_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.0000001 --max_lr 0.00001 --save_freq 20 --epochs 81 --resume 'results/dl/unet_z234_12bs_4augs_ft_80epochs/checkpoints/final.pth'
```

_Zone 2, 4 (For zone 4)_
```
python hrlcm/fit_compose_lr.py --exp_name unet_z24_12bs_4augs_ft_fz56_80epochs --num_workers 48 --gpu_devices '0,1,2,3' --min_lr 0.0000001 --max_lr 0.00001 --save_freq 20 --epochs 81 --resume 'results/dl/unet_z234_12bs_4augs_ft_80epochs/checkpoints/final.pth'
```

@@

Doing prediction is simple. Applying a model into its zones.

@@

For example:
```
python hrlcm/predict.py --args_path 'results/dl/unet_z125_ft_80epochs/checkpoints/args.pkl' --checkpoint_file 'results/dl/unet_z125_ft_80epochs/checkpoints/final.pth' --fname_predict 'dl_catalog_predict.csv' --data_dir 'results/tanzania' --out_dir 'results/prediction' --stats_dir 'results/tanzania/norm_stats' --num_workers 48 --batch_size 32 --gpu_devices '0,1,2,3'

aws s3 cp --recursive $HOME/hrlcm/results/prediction s3://activemapper/leisong/prediction
```

Steps in production:

```

python hrlcm/predict.py --args_path 'results/dl/unet_z125_ft_80epochs/checkpoints/args.pkl' --checkpoint_file 'results/dl/unet_z125_ft_80epochs/checkpoints/final.pth' --fname_predict 'dl_catalog_predict_zone1_1.csv' --data_dir 'results/tanzania' --out_dir 'results/prediction' --stats_dir 'results/tanzania/norm_stats' --num_workers 48 --batch_size 32 --gpu_devices '0,1,2,3'

aws s3 cp --recursive $HOME/hrlcm/results/prediction s3://activemapper/leisong/prediction/zone1

rm -rf results/prediction
```

```{r}
# Get imgs in S3
library(aws.s3)
library(dplyr)
library(parallel)
items <- get_bucket(bucket = "activemapper", 
                    prefix = "leisong/predict", 
                    max = Inf)
keys <- unlist(lapply(c(2: length(items)), function(i) {
    gsub("leisong/", "", items[[i]]$Key)
}))

catalog_pred <- read.csv("results/tanzania/dl_catalog_predict.csv", 
                         stringsAsFactors = FALSE)
fns <- catalog_pred$img %>% unique()

fns <- setdiff(fns, keys)
pth_from <- '/Volumes/elephant'
pth_to <- 's3://activemapper/leisong'
mclapply(fns, function(tile_pth) {
    print(tile_pth)
    cmd <- sprintf('aws s3 cp %s/%s %s/%s',
                   pth_from, tile_pth, pth_to, tile_pth)
    system(cmd, intern = TRUE)
}, mc.cores = 10)
```

@@

### Copy files from/to S3

To:

```
aws s3 cp --recursive $HOME/hrlcm/results/dl s3://activemapper/leisong/model

while true; do aws s3 cp --recursive $HOME/hrlcm/results/dl/unet_full_startup_80epochs s3://activemapper/leisong/model/unet_full_startup_80epochs; sleep 50m; done

aws s3 cp --recursive $HOME/hrlcm/results/prediction_zone1_125 s3://activemapper/leisong/prediction/
```

From:

```
aws s3 cp --recursive s3://activemapper/leisong/unet_full_12bs_4augs_120epochs/ $HOME/hrlcm/results/dl/unet_full_12bs_4augs_120epochs
```
